Let $R$ be a ring.
We say that a function $f : R \to R$ is \emph{good (over $R$)} if for any $x, y \in R$,
\[ f(f(x) f(y)) + f(x + y) = f(xy). \tag{*}\label{2017a6-eq0} \]
Find all good functions over any ring.





\section{Main Results}

Unfortunately, we still don't know the complete answer to the above general question.
However, we have one main conjecture, saying what would the answer be.
For any ring $R$, double-sided ideal $I$, and $x \in R$, we denote by $[x]$ the image of $x$ in $R/I$ whenever the context is clear.
The main conjecture is as follows.

\begin{conjecture}\label{2017a6-s1-conj-main}
Let $R$ be a ring.
Consider triples $(I, \iota, a)$, where
\begin{itemize}
    \item   $I$ is a double-sided ideal of $R$;
    \item   $\iota : R/I \to R$ is a homomorphism of additive group;
    \item   $\iota$ is a right inverse to the natural projection map $R \to R/I$, i.e. $[\iota(y)] = y$ for any $y \in R/I$;
    \item   $a$ is an element of $Z(R/I)$ such that $a^2 = 1$.
\end{itemize}
Then the good functions over $R$ are precisely those of form $f(x) = \iota(a(1 - [x]))$ for some triple $(I, \iota, a)$.
\end{conjecture}

We prove the conjecture in some subcases.
The results we prove are as follows.

\begin{theorem}\label{2017a6-s1-general}
Conjecture~\ref{2017a6-s1-conj-main} holds when $2$ is a unit in $R$ and $3$ is not a zero divisor in $R$.
\end{theorem}

\begin{theorem}\label{2017a6-s1-simple-ring-char-ne-two}
Conjecture~\ref{2017a6-s1-conj-main} holds when $R$ is a simple ring with $\rchar(R) \neq 2$.
That is, for any simple ring $R$ of characteristic $\neq 2$, the only good functions over $R$ are $0$, $x \mapsto x - 1$, and $x \mapsto 1 - x$.
\end{theorem}

\begin{theorem}\label{2017a6-s1-field-char-two}
Conjecture~\ref{2017a6-s1-conj-main} holds when $R$ is a field with $\rchar(R) = 2$.
That is, for any field $F$ of characteristic $2$, the only good functions on $F$ are $0$ and $x \mapsto x + 1$.
\end{theorem}

As usual, it is relatively easier to verify if a function is good with sufficient explicit description.
The bulk of the work is to prove that no other functions are good.
To check that $f(x) = \iota(a(1 - [x]))$ is good, we just plug in directly to~\eqref{2017a6-eq0}.
First note that since $\phi$ is a ring homomorphism, for any $x, y \in R$,
\[ [f(x) f(y)] = [\iota(a(1 - [x]))] [\iota(a(1 - [y]))] = a(1 - [x]) a(1 - [y]). \]
Since $a \in Z(R/I)$ and $a^2 = 1$, the right hand side is equal to $(1 - [x])(1 - [y])$.
This implies
\[ f(f(x) f(y)) = \iota(a(1 - (1 - [x])(1 - [y]))) = \iota(a[x + y - xy]) = \iota(a(1 - [xy])) - \iota(a(1 - [x + y])), \]
    where the last equality holds since $\iota$ is additive.
But the right hand side is $f(xy) - f(x + y)$ by definition, so we are done.
It now remains to prove that these are all the good functions.

Some functions that obviously work are $x \mapsto 0$, $x \mapsto 1 - x$, $x \mapsto 0$, and $x \mapsto x - 1$.
The following example shows a more convoluted good function and also illustrate why Conjecture~\ref{2017a6-s1-conj-main} cannot be simplified.

\begin{example}
Let $R$ be a ring.
Define $f : R[X] \to R[X]$ by $f(P) = (1 - P(1)) X^k$ for all $P \in R[X]$, where $k$ is an arbitrary non-negative integer.
Then one can check by hand that $f$ is a good function over $R$.
This example corresponds to $S = R$, with the projection $\phi$ defined mod $X - 1$, the lift $\iota$ defined by $P \mapsto X^k P$, and $a = 1$.    
\end{example}





\section{General Steps of Proof}

We now describe the major general steps we use to attempt proving Conjecture~\ref{2017a6-s1-conj-main}.
In the original problem, $f(1) = 0$ is easy to obtain.
This is not the case in general, but we will still prove it.
The method we use to prove $f(1) = 0$ tells us something more about good functions in general.

We start by obtaining basic equalities.
First, plugging $x = y = 0$ into~\eqref{2017a6-eq0} yields $f(f(0)^2) = 0$.
Plugging $y = 1$ into~\eqref{2017a6-eq0} yields $f(x + 1) = f(x) - f(f(x) f(1))$ for any $x \in R$.
In particular, this yields that for any $a, b \in R$,
\[ f(a) = f(b) \implies f(a + 1) = f(b + 1). \tag{2.1}\label{2017a6-eq2-1} \]

We now prove some main results.
Denote the set of periods of $f$ by
\[ I_f := \{c \in R : \forall x \in R, f(x + c) = f(x)\}. \tag{2.2}\label{2017a6-eq2-2} \]
In general, if $I$ is a double-sided ideal contained in $I_f$, then we obtain an induced function $\tilde{f} : R/I \to R/I$ defined by
\[ \tilde{f}(x + I) = f(x) + I \quad \forall x \in R. \]
Then one can check by hand that $\tilde{f}$ is in fact a good function.
Furthermore, for any $c \in R$, we have $c + I \in I_{\tilde{f}}$ if and only if $c \in I_f$.
This means that $\tilde{f}$ potentially has one additional nice properties compared to $f$.
It would be nice if $I_f$ is already a double-sided ideal.
The following theorem says that it is indeed the case.

\begin{theorem}\label{2017a6-s2-period-quot}
The set $I_f$ is in fact a double-sided ideal.
The induced function from $R/I_f$ to itself defined by $[x] \mapsto [f(x)]$ is good with no non-zero period.
\end{theorem}
\begin{proof}
By the previous discussion, it only remains to prove that $I_f$ is a double-sided ideal.
Clearly $I_f$ is closed under addition and additive inverse.
Now fix any $c \in I_f$.
Comparing~\eqref{2017a6-eq0} using $y = c$ and $y = 0$ gives $f(xc) = f(0)$ for all $x \in R$.
Then replacing $y$ with $yc$ yields
\[ f(x) + f(f(x) f(0)) = f(0) = f(xyc) = f(x + yc) + f(f(x) f(yc)) = f(x + yc) + f(f(x) f(0)), \]
    so $f(x + yc) = f(x)$ for all $x, y \in R$.
That is, $yc \in I_f$ for any $y \in R$.
Similarly, we also get $cy \in I_f$ for any $y \in R$.
This shows that $I_f$ is a double-sided ideal.
\end{proof}

We can now show that $f(1) = 0$.
More generally, we prove:

\begin{theorem}\label{2017a6-s2-preimage-zero}
For any $C \in R$, if $f(C) = 0$, then $C \equiv 1 \pmod{I_f}$.
\end{theorem}
\begin{proof}
From~\eqref{2017a6-eq0}, we have $f(Cx) = f(0) + f(C + x)$ for all $x \in R$.
Then we get $f(C + 1) = -f(0)$ and $f(-C^2) = 2 f(0)$.
Plugging $x = 0$ and $y = -C^2$ into~\eqref{2017a6-eq0} yields $f(2 f(0)^2) = -f(0)$.

On the other hand, note that $f(C) = f(f(0)^2) = 0$.
Applying~\eqref{2017a6-eq2-1} twice yields $f(C + 2) = f(f(0)^2 + 2)$.
Now plug in~\eqref{2017a6-eq0} with $y = 2$ and with $x = C$ vs. $x = f(0)^2$.
This yields $f(2C) = f(2 f(0)^2) = -f(0)$.

Plugging $x = y = C$ into~\eqref{2017a6-eq0} yields $f(C^2) = 0$.
Plugging $x = C$ and $y = C^2$ yields $f(0) + f(C^2 + C) = f(C^3)$.
On the other hand, since $f(C) = 0$, we get $f(C^2 + C) = f(2C + 1) + f(0)$.
Since $f(C + 1) = f(2C) = -f(0)$,~\eqref{2017a6-eq2-1} yields
\[ f(2C + 1) = f(C + 2) = f(2C) - f(0) = -2 f(0). \]
By~\eqref{2017a6-eq0} with $x = C$ and $y = C + 1$, we get $f(C^2 + C) = -f(0)$.
By~\eqref{2017a6-eq0} with $x = C$ and $y = C^2$, we get $f(C^3) = 0$.

Now we are ready for the final step.
For any $x \in R$, we write $f(C^4 x)$ in two ways, using the fact that $f(C) = f(C^2) = f(C^3) = 0$.
\[ f(C^4 x) = f(C^2 + C^2 x) + f(0) = f(C^2 (x + 1)) + f(0) = f(C^2 + x + 1) + 2 f(0), \]
\[ f(C^4 x) = f(C^3 + Cx) + f(0) = f(C (C^2 + x)) + f(0) = f(C^2 + x + C) + 2 f(0). \]
Replacing $x$ with $x - C^2$ gives $f(x + C) = f(x + 1)$ for any $x \in R$.
This proves that $C \equiv 1 \pmod{I_f}$.
\end{proof}

Since $f(f(0)^2) = 0$, Theorem~\ref{2017a6-s2-preimage-zero} yields $f(0)^2 \equiv 1 \pmod{I_f}$.
Thus $f(1) = f(f(0)^2) = 0$.
We prove some more properties that allow further reductions.

Plugging $y = 1$ into~\eqref{2017a6-eq0} now yields $f(x + 1) = f(x) - f(0)$ for all $x \in R$.
Next,~\eqref{2017a6-eq0} with $y = 0$ or $x = 0$ yields
\[ f(f(x) f(0)) = f(f(0) f(x)) = f(0) - f(x) \quad \forall x \in R. \tag{2.3}\label{2017a6-eq2-3} \]
Replacing $x$ with $f(x) f(0)$ in~\eqref{2017a6-eq2-3} yields
\[ f(x) = f((f(0) - f(x)) f(0)) = f(1 - f(x) f(0)) = f(-f(x) f(0)) - f(0). \]
Thus $f(-f(x) f(0)) = f(x) + f(0) = f(x - 1)$.
On the other hand, plugging $y = -1$ yields
\[ f(f(x) f(-1)) + f(x) + f(0) = f(-x) \quad \forall x \in R. \]
In particular, for any $x, y \in R$ such that $f(x) = f(y)$, we have $f(-x) = f(-y)$.
Thus the previous equality and~\eqref{2017a6-eq2-3} yields
\[ f(f(x) f(0)) = f(-(x - 1)) = f(1 - x). \tag{2.4}\label{2017a6-eq2-4} \]
By~\eqref{2017a6-eq2-3}, we get
\[ f(x) + f(-x) = 2 f(0). \tag{2.5}\label{2017a6-eq2-5} \]

\begin{theorem}\label{2017a6-s2-map-zero-reduction}
Suppose that $f : R \to R$ is good with $I_f = \{0\}$.
Then $f(0) f(x) = f(x) f(0)$ for any $x \in R$.
The function $g : R \to R$ defined by $g(x) = f(0) f(x)$ is good with $I_g = \{0\}$ and $g(0) = 1$.
\end{theorem}
\begin{proof}
For the first sentence, we first substitute $(f(x) f(0), f(0) f(1 - x))$ for $(x, y)$ in~\eqref{2017a6-eq0}.
Since $I_f = \{0\}$, we get $f(0)^2 = 1$.
Applying~\eqref{2017a6-eq2-4} gives
\[ f(f(1 - x) f(x)) + f(f(x) f(0) + f(0) f(1 - x)) = f(f(x) f(0)^2 f(1 - x)) = f(f(x) f(1 - x)). \]
Now apply~\eqref{2017a6-eq0} on two of the terms:
\[ f((1 - x) x) - f(1) + f(f(x) f(0) + f(0) f(1 - x)) = f(x (1 - x)) - f(1). \]
Since $(1 - x)x = x(1 - x)$, we get $f(f(x) f(0) + f(0) f(1 - x)) = 0$.
By Theorem~\ref{2017a6-s2-preimage-zero} and $I_f = \{0\}$,
\[ f(x) f(0) + f(0) f(1 - x) = 1. \]
By~\eqref{2017a6-eq2-5} and $f(x + 1) = f(x) - f(0)$, we get $f(1 - x) = f(0) - f(x)$.
Since $f(0)^2 = 1$, this yields $f(0) f(1 - x) = 1 - f(0) f(x)$.
Thus we get $f(x) f(0) = f(0) f(x)$, as desired.

Now we prove the second sentence.
Clearly $g(0) = 1$, so we just check that $g$ is good with $I_g = \{0\}$.
Indeed, we have
\[ g(g(x) g(y)) = f(0) f(f(0) f(x) f(0) f(y)) = f(0) f(f(x) f(y)). \]
By~\eqref{2017a6-eq0}, $f(0) f(f(x) f(y)) = f(0) f(xy) - f(0) f(x + y) = g(xy) - g(x + y)$.
This proves that $g$ is good.
Finally, it is straightforward to check that $I_g = I_f = \{0\}$.
\end{proof}



From now on, we say that a good function $f : R \to R$ is \emph{reduced good} if $I_f = \{0\}$ and $f(0) = 1$.
By Theorem~\ref{2017a6-s2-preimage-zero}, $I_f = \{0\}$ is equivalent to saying that $f(C) = 0$ if and only if $C = 1$.
To summarize the progress above, we have the following results:

\begin{theorem}\label{2017a6-s2-good-to-reduced-good}
Let $f : R \to R$ be a good function, and define $\tilde{f} : R/I_f \to R$ by $\tilde{f}([x]) = f(x)$ for all $x \in R$.
Then there exists $a \in Z(R/I_f)$ with $a^2 = 1$ and a reduced good function $h : R/I_f \to R/I_f$ such that
\begin{itemize}
    \item   $[\tilde{f}(y)] = a h(y)$ for any $y \in R/I_f$;
    \item   $\tilde{f}(h(x) h(y)) + \tilde{f}(x + y) = \tilde{f}(xy)$ for all $x, y \in R/I_f$.
\end{itemize}
\end{theorem}
\begin{proof}
Take $a = [f(0)] = [\tilde{f}(0)]$, which satisfies $[f(0)]^2 = 1$ by Theorem~\ref{2017a6-s2-preimage-zero}.
The map $y \mapsto [\tilde{f}(y)]$ is good with no non-zero period by Theorem~\ref{2017a6-s2-period-quot}.
Thus, the function $h(y) = a [\tilde{f}(y)]$ is reduced good by Theorem~\ref{2017a6-s2-map-zero-reduction}.
It also satisfies $a h(y) = a^2 [\tilde{f}(y)] = [\tilde{f}(y)]$ for any $y \in R/I_f$.
Finally,~\eqref{2017a6-eq0} yields that for any $x, y \in R$,
\[ \tilde{f}([f(x) f(y)]) + \tilde{f}([x] + [y]) = \tilde{f}([x][y]). \]
However, \[ [f(x) f(y)] = [\tilde{f}([x]) \tilde{f}([y])] = [\tilde{f}([x])] a^2 [\tilde{f}([y])]
    = a [\tilde{f}([x])] a [\tilde{f}([y])] = h([x]) h([y]). \]
This proves the second bullet; we are done.
\end{proof}

\begin{theorem}\label{2017a6-s2-reduced-good-prop}
Let $f : R \to R$ be a reduced good function.
Then $f(x + 1) = f(x) - 1$, $f(f(x)) = 1 - f(x)$, and $f(-x) = 2 - f(x)$ for any $x \in R$.
\end{theorem}
\begin{proof}
For the first equality, plug $y = 1$ and note that $f(1) = 0$ and $f(0) = 1$.
The other two equalities follow from~\eqref{2017a6-eq2-3} and~\eqref{2017a6-eq2-5}.
\end{proof}

We close this section with some conjectures that, if true, proves Theorem~\ref{2017a6-s1-conj-main}.
Ideally, we hope that the conjecture is true.
Later, we prove these conjectures in some subcases.

\begin{conjecture}\label{2017a6-s2-conj1}
The only reduced good function over a fixed ring is $x \mapsto 1 - x$.
\end{conjecture}

\begin{conjecture}\label{2017a6-s2-conj2}
Let $R$ be a ring and $(G, +)$ be an abelian group.
Let $f : R \to G$ be a function such that for any $x, y \in R$,
\[ f((1 - x)(1 - y)) + f(x + y) = f(xy). \]
Then the function $g(x) = f(1 - x)$ is additive, i.e. a group homomorphism.
\end{conjecture}

We now see why these conjectures imply the desired result.

\begin{theorem}
Conjecture~\ref{2017a6-s2-conj1} and Conjecture~\ref{2017a6-s2-conj2} implies Conjecture~\ref{2017a6-s1-conj-main}.
\end{theorem}
\begin{proof}
Let $f$ be a good function.
Let $\tilde{f}$, $a$, and $h$ be as in Theorem~\ref{2017a6-s2-good-to-reduced-good}.
We get the desired $a$, and we can take $I = I_f$.
So, it remains to find $\iota$ and the formula for $f$ in terms of $\iota$ and $a$.

Conjecture~\ref{2017a6-s2-conj1} yields $h(x) = 1 - x$ for all $x \in R$.
Then Conjecture~\ref{2017a6-s2-conj2} would imply that $x \mapsto \tilde{f}(1 - x)$ is a group homomorphism.
So we can write $\tilde{f}(x) = \iota'(1 - x)$ for some group homomorphism $\iota' : R/I_f \to R$.
Then we get the formula
\[ f(x) = \tilde{f}([x]) = \iota'(1 - [x]). \]

Now take $\iota(x) = \iota'(ax)$.
Then $a^2 = 1$ yields $\iota'(x) = \iota(ax)$ and $f(x) = \iota(a(1 - [x]))$ for all $x \in R$.
Finally, recall from Theorem~\ref{2017a6-s2-good-to-reduced-good} that $[\tilde{f}(x)] = a h(x) = a(1 - x)$ for all $x \in R/I_f$.
Thus we get
\[ [\iota(x)] = [\iota'(ax)] = [\tilde{f}(1 - ax)] = [a(1 - (1 - ax))] = [a^2 x] = [x]. \]
This proves the theorem.
\end{proof}





\section{Progress towards Conjecture~\ref{2017a6-s2-conj1}}

In this section, we solve Conjecture~\ref{2017a6-s2-conj1} partially.
That is, we characterize reduced good functions.
We take the three equations from Theorem~\ref{2017a6-s2-reduced-good-prop}.
Here we are also able to prove Theorem~\ref{2017a6-s1-simple-ring-char-ne-two} and Theorem~\ref{2017a6-s1-field-char-two}.

\begin{theorem}\label{2017a6-s3-main1}
Let $R$ be a ring such that $2$ is not a zero divisor in $R$.
Then the only reduced good functions over $R$ is $x \mapsto 1 - x$.
\end{theorem}
\begin{proof}
From $f(f(x)) = 1 - f(x)$, we have $f(f(f(x))) = f(x)$ for all $x \in R$.
From the two equalities, it remains to show that $f$ is injective.

Let $a, b \in R$ such that $f(a) = f(b)$.
Then $f(f(a) f(b)) = f(f(b) f(a))$, so~\eqref{2017a6-eq0} yields $f(ab) = f(ba)$.
Next, $f(-a) = 2 - f(a) = 2 - f(b) = f(-b)$, so $f(f(-a) f(b)) = f(f(-b) f(a))$.
Then~\eqref{2017a6-eq0} yields $f(-ab) - f(b - a) = f(-ba) - f(a - b)$.
But we've seen that $f(ab) = f(ba)$, so $f(-ab) = f(-ba)$ and thus $f(b - a) = f(a - b)$.
By the equation $f(-x) = 2 - f(x)$, we get $2 f(a - b) = 2$.
Since $2$ is not a zero divisor in $R$, this yields
\[ f(a - b) = 1 \implies f(a - b + 1) = 1 - 1 = 0 \implies a - b + 1 = 1 \implies a = b. \]
This proves that $f$ is injective; we are done.
\end{proof}

\begin{proof}[Proof of Theorem~\ref{2017a6-s1-simple-ring-char-ne-two}]
Let $R$ be a simple ring with $\rchar(R) \neq 2$, and let $f : R \to R$ be a good function.
The only two-sided ideals of $R$ are $\{0\}$ and $R$, so $I_f \in \{\{0\}, R\}$.
If $I_f = R$, then clearly $f \equiv 0$.
So now, assume that $I_f = \{0\}$.

By Theorem~\ref{2017a6-s2-map-zero-reduction}, $f(0)$ commutes with $f(x)$ for all $x \in R$.
Also, the function $x \mapsto f(0) f(x)$ is reduced good.
Then Theorem~\ref{2017a6-s3-main1} implies that $f(0) f(x) = 1 - x$ for all $x \in R$.
Since Theorem~\ref{2017a6-s2-preimage-zero} implies $f(0)^2 = 1$, we get $f(x) = f(0)(1 - x)$ for all $x \in R$.
In particular, $f$ is bijetive and $f(0) \in Z(R)$.
Since $R$ is simple, this and $f(0)^2 = 1$ yields $f(0) = \pm 1$.
We are done.
\end{proof}

To prove Theorem~\ref{2017a6-s1-field-char-two}, we focus on what happens with the units of $R$.

\begin{theorem}\label{2017a6-s3-main2}
Let $R$ be an integral domain of characteristic $2$ and $f : R \to R$ be a reduced good function.
Then $f(x) = x + 1$ for any $x \in R^{\times}$.
\end{theorem}
\begin{proof}
First note that if $x \in R^{\times}$, then
\[ f(f(x + 1) f(x^{-1} + 1)) = f((x + 1)(x^{-1} + 1)) - f(x + x^{-1} + 2) = 0 \iff f(x + 1) f(x^{-1} + 1) = 1. \]
Similarly $f(x^{-1} + 1) f(x + 1) = 1$.
Thus $f(x + 1) \in R^{\times}$ as well and $f(x^{-1} + 1) = f(x + 1)^{-1}$.
Notice that $f(f(x + 1)) = 1 - f(x + 1) = f(-x)$ and both $f(x + 1)$ and $-x$ are units of $R$.
It remains to show the following: if $a, b \in R^{\times}$ and $f(a) = f(b)$, then $a = b$.

Consider~\eqref{2017a6-eq0} with $(x, y) = (a + 1, b^{-1} + 1)$.
Since $f(a + 1) = f(b + 1)$, we get
\begin{align*}
    f((a + 1)(b^{-1} + 1)) &= f((a + 1) + (b^{-1} + 1)) \\
    f(ab^{-1} + a + b^{-1} + 1) &= f(a + b^{-1}) \\
    f(ab^{-1} + a + b^{-1}) &= f(1 + a + b^{-1})
\end{align*}
Now notice the identity
\[ (1 + a + b^{-1})(1 + b + a^{-1}) = (ab^{-1} + a + b^{-1})(ba^{-1} + b + a^{-1}). \]
As a result, by plugging the appropriate values into~\eqref{2017a6-eq0}, we get
\[ f((1 + a + b^{-1}) + (1 + b + a^{-1})) = f((ab^{-1} + a + b^{-1}) + (ba^{-1} + b + a^{-1})). \]
Letting $C = a + b + 1$ and $D = a^{-1} + b^{-1} + 1$, the above equation is equivalent to saying that
\[ f(C + D) = f(CD + 1) = f(CD) + 1. \]
But then plugging $(x, y) = (C, D)$ into~\eqref{2017a6-eq0} yields
\[ f(f(C) f(D)) = 1 \iff f(C) f(D) = 0 \iff f(C) = 0 \vee f(D) = 0 \iff C = 1 \vee D = 1. \]
By definition of $C$ and $D$, and by $\rchar(R) = 2$, this is equivalent to
\[ a + b + 1 = 1 \vee a^{-1} + b^{-1} + 1 = 1 \iff a = b \vee a^{-1} = b^{-1}. \]
But $a^{-1} = b^{-1}$ yields $a = b$.
So, regardless, we have obtained $a = b$.
This proves that $f$ is injective.
\end{proof}

\begin{proof}[Proof of Theorem~\ref{2017a6-s1-field-char-two}]
Let $F$ be a field of characteristic $2$ and $f : F \to F$ be a good function.
Then $I_f \in \{\{0\}, F\}$, and $I_f = F$ yields $f \equiv 0$.
Otherwise, if $I_f = \{0\}$, then $f(0)^2 = 1 \iff f(0) = \pm 1$.
Since $\rchar(F) = 2$, this gives $f(0) = 1$, so $f$ is reduced good.
By Theorem~\ref{2017a6-s3-main2}, $f(x) = x + 1$ for all $x \neq 0$.
Since $f(0) = 1$, we get $f(x) = x + 1$ for all $x \in F$, as desired.
\end{proof}





\section{Progress towards Conjecture~\ref{2017a6-s2-conj2}}

Here we write down some progress towards proving Conjecture~\ref{2017a6-s2-conj2}.
This, along with Theorem~\ref{2017a6-s3-main1}, proves Theorem~\ref{2017a6-s1-general}.

\begin{theorem}\label{2017a6-s4-main}
Let $R$ be a ring and $(G, +)$ be an abelian group.
Let $f : R \to G$ be a function such that for any $x, y \in R$,
\[ f((1 - x)(1 - y)) + f(x + y) = f(xy). \]
Then $h(x) = 6 f(1 - x)$ is a group homomorphism.
\end{theorem}
\begin{proof}
Plugging $x = y = 0$ yields $g(0) = 0$.
Next, plugging $y = 1$ and replacing $x$ with $-x$ yields $g(x + 1) = g(x) + g(1)$.
By induction on $n \in \N$, we get
\[ g(x + n) = g(x) + n g(1) \quad \forall x \in R, n \in \N. \tag{4.1}\label{2017a6-s4-main-eq1} \]
Plugging $y = 0$ and then using~\eqref{2017a6-s4-main-eq1} yields
\[ g(-x) = -g(x) \quad \forall x \in R. \tag{4.2}\label{2017a6-s4-main-eq2} \]
After some rearrangement,~\eqref{2017a6-s4-main-eq1} now yields
\[ g((x + 1)(y + 1)) = g(xy) + g(x + y) + g(1) \quad \forall x, y \in R. \]
By induction on $n \in \N$,
\[ g((x + n)(y + n)) = g(xy) + n g(x + y) + n^2 g(1) \quad \forall x, y \in R, n \in \N. \tag{4.3}\label{2017a6-s4-main-eq3} \]
Plugging $y = 0$ and replacing $x$ with $x - n$ yields
\[ g(nx) = n g(x) \quad \forall x \in R, n \in \N. \tag{4.4}\label{2017a6-s4-main-eq4} \]

Now consider that by~\eqref{2017a6-s4-main-eq3}, for any $x, y \in R$,
\begin{align*}
    g((x + 2)(2y + 2)) &= g(2xy) + 2 g(x + 2y) + 4 g(1) \\
    2 g((x + 2)(y + 1)) &= 2 (g(xy) + g(x + 2y) + 2 g(1)).
\end{align*}
Similarly, for any $x, y \in R$,
\[ 2 g((x + 1)(y + 2)) = 2 (g(xy) + g(2x + y) + 2 g(1)). \]
Combining the two, it can be shown that
\[ 2 g((x + 3)(y + 3)) = 2 (g(xy) + g(x + 2y) + g(2x + y) + 9 g(1)). \]
Then~\eqref{2017a6-s4-main-eq3} yields
\[ 2 (g(x + 2y) + g(2x + y)) = 6 g(x + y) \quad \forall x, y \in R. \]
Replace $x$ and $y$ with $2x - y$ and $2y - x$, and we get
\[ 6 (g(x) + g(y)) = 6 g(x + y). \]
Since $h(x) = 6 f(1 - x) = 6 g(x)$ for all $x \in R$, we get the desired result.
\end{proof}

\begin{corollary}\label{2017a6-s4-23-nzd}
Let $R$ be a ring such that $2$ and $3$ are not zero divisors of $R$ and let $(G, +)$ be an abelian group.
Let $f : R \to G$ be a function such that for any $x, y \in R$,
\[ f((1 - x)(1 - y)) + f(x + y) = f(xy). \]
Then $g(x) = f(1 - x)$ is a group homomorphism.
\end{corollary}

\begin{proof}[Proof of Theorem~\ref{2017a6-s1-general}]
Let $R$ be a ring, and first assume only that $2$ is a unit.
Then $2$ is not a zero divisor in $R/I_f$ no matter what the ideal $I_f$ is.
By Theorem~\ref{2017a6-s3-main1}, the only reduced good function over $R$ is $x \mapsto 1 - x$.
By Theorem~\ref{2017a6-s2-good-to-reduced-good}, there exists $a \in Z(R/I_f)$ with $a^2 = 1$ such that
\begin{itemize}
    \item   $[\tilde{f}(y)] = a(1 - y)$ for any $y \in R/I_f$;
    \item   $\tilde{f}((1 - x)(1 - y)) + \tilde{f}(x + y) = \tilde{f}(xy)$ for any $x, y \in R/I_f$.
\end{itemize}

Now assume that $3$ is not a zero divisor in $R$.
Then by the above corollary, the map $\iota' : R/I_f \to R$ defined by $\iota'(x) = \tilde{f}(1 - x)$ is a group homomorphism.
By direct computation, $[\iota'(y)] = ay$ for all $y \in R/I_f$.
Setting $\iota(y) = \iota'(ay)$. we get that $\iota$ is a homomorphism of additive group and $[\iota(y)] = y$ for all $y \in R/I_f$.
Finally, we compute and get $f(x) = \tilde{f}([x]) = \iota'(1 - [x]) = \iota(a(1 - [x]))$ for all $x \in R$.
\end{proof}





\section*{Further directions}

The set of functions $f : R \to G$ satisfying $f((1 - x)(1 - y)) + f(x + y) = f(xy)$ is a group under addition, say $H(R, G)$.
It contains $\Hom(R, G)$, so let's consider $H_0(R, G) = H(R, G)/\Hom(R, G)$.
If $R$ surjects to $S$, then $H(S, G)$ injects to $H(R, G)$ and $H_0(S, G)$ injects to $H_0(R, G)$.
That means if you want to show that $H_0(R, G) = 1$ for all $R$ and $G$, it suffices to consider the case where $R$ is a free $\Z$-algebra.

Suppose that $R = \Z \langle \Sigma \rangle$ for some set $\Sigma$.
Then $H_0(R, G)$ is isomorphic to a subset of $H(R, G)$ consisting of the functions $f$ taking value zero on the monomials.
We proved a result saying that $H_0(R, G)$ is $6$-torsion in general.
Thus we can split it into the $2$-torsion part $H_{0, 2}(R, G)$ and the $3$-torsion part $H_{0, 3}(R, G)$.
Back to the monomial case, the functions in $H_{0, k}(R, G)$ are $k$-torsion for $k = 2, 3$.
Thus we can assume that $G$ is an $\F_k$-vector space.
Furthermore, we can quotient down $R$ and assume that $R$ is a free $\F_3$-algebra for $k = 3$.

If $R = R' \oplus M$ for some $R$-module $M$ (multiplication on $M$ is defined by $m_1 m_2 = 0$), then one can show that $H_0(R, G) = H_0(R', G)$.
If $R = R_1 \times R_2$, then $H_0(R, G) \cong H_0(R_1, G) \times H_0(R_2, G)$.
I think this also generalizes to infinite product.

One possible way to approach the problem: "find a triangle".
That is, for any $a_1, a_2, a_3 \in R$, find $x_1, x_2, x_3, y_1, y_2, y_3 \in R$ such that:
\[ a_i = x_i y_i = (x_{i + 1} + 1)(y_{i + 1} + 1) - 1, \quad \forall i \leq 3. \]
Well, this notation assumes $\rchar(R) = 2$, I think.
But this should be adaptable to deal with the characteristic $3$ case.
