A multiset $S$ of positive rational numbers is called \emph{nice} if $\sum_{q \in S} q$ and $\sum_{q \in S} 1/q$ are both integers.
Find all $n \in \N$ such that there exists infinitely many nice multisets $S$ of size $n$.



\subsection*{Answer}

$n \geq 3$.



\subsection*{Solution}

Official solution: \url{http://www.imo-official.org/problems/IMO2017SL.pdf}

We follow Solution 1 of the official solution.
However, the constructions for $n = 3$ will be explicit; it is described in the comment section.

Clearly, by adding occurrences of the integer $1$ in $S$, if $n \geq 0$ is good, then $n + 1$ is good.
Thus, it suffices to show that $2$ is not good but $3$ is good.

\begin{claim}
There are only three nice multisets of size $2$.
Thus, $2$ is not good.
\end{claim}
\begin{proof}
A nice multiset of size $2$ corresponds to a pair $(x, y) \in \Q^+$ such that $x + y$ and $1/x + 1/y$ are integers.
Write $x = a/b$ and $y = c/d$, where $a, b, c, d \in \N^+$ and $\gcd(a, b) = \gcd(c, d) = 1$.
Just from $x + y \in \Z$, we claim that $b = d$.
Indeed, we get $bd \mid ad + bc$, so $b \mid ad$ and thus $b \mid d$ since $\gcd(a, b) = 1$.
Similarly, $d \mid b$, so $b = d$.

From $1/x + 1/y \in \Z$, we also get $a = c$.
In particular, this means $x = y$ and $2x, 2/x \in \Z$.
Direct bashing yields $x = y \in \{1/2, 1, 2\}$, proving the claim.
\end{proof}

\begin{claim}
For each $n \geq 0$, let $F_n$ denote the $n^{\text{th}}$ Fibonacci number.\footnote{By convention, $F_0 = 0$ and $F_1 = 1$.}
Fix $k \geq 0$, and define $A_k = F_{2k + 3} + F_{2k + 1} + 3$.
Then the following multiset is nice:
\[ T_k := \left\{\frac{1}{A_k}, \frac{F_{2k + 1} + 1}{A_k}, \frac{F_{2k + 3} + 1}{A_k}\right\}. \]
As a result, $3$ is good.
\end{claim}
\begin{proof}
The sum of the elements of $T_k$ is $1$, so it remains to prove that the sum of their reciprocals is an integer.
Indeed, we have
\[ A_k + \frac{A_k}{F_{2k + 1}} + \frac{A_k}{F_{2k + 3}} = (A_k + 2) + \frac{F_{2k + 3} + 2}{F_{2k + 1} + 1} + \frac{F_{2k + 1} + 2}{F_{2k + 3} + 1}; \]
We claim that
\[ \frac{F_{2k + 3} + 2}{F_{2k + 1} + 1} + \frac{F_{2k + 1} + 2}{F_{2k + 3} + 1} = 3. \]
This can be rewritten as
\[ F_{2k + 3}^2 + F_{2k + 1}^2 + 1 = 3 F_{2k + 3} F_{2k + 1}, \]
    or just $F_{2k + 2}^2 + 1 = F_{2k + 3} F_{2k + 1}$.
This is a well-known identity for Fibonacci numbers.
More generally, one can show by induction that $F_{n + 1}^2 = F_{n + 2} F_n + (-1)^n$ for all $n \geq 0$.
\end{proof}
